### 2024-01-15 TOD4.qmd
Add files via upload

---
title: "TRT Dinesafe Data"
format: revealjs
editor: visual
author: Tianen (Evan) Hao
format: pdf
bibliography: bibliography.bib
---

```{r}
install.packages("readr")

```

```{r}
library(readr)
library(dplyr)
library(ggplot2)

data_path <- "/cloud/project/data/Dinesafe.csv"
dataset <- read_csv(data_path)

cleaned_data <- dataset %>%
  filter(!is.na(`Establishment Name`)) %>%
  mutate(new_column = tolower(`Establishment Name`))

```

```{r}
install.packages("AER")
install.packages("future")
install.packages("gitcreds")
install.packages("knitr")
install.packages("lintr")
install.packages("renv")
install.packages("reprex")
install.packages("styler")
install.packages("tictoc")
install.packages("tidyverse")
install.packages("tinytex")
install.packages("usethis")

```

```{r}
library(AER)
library(future)
library(gitcreds)
library(knitr)
library(lintr)
library(renv)
library(reprex)
library(styler)
library(tictoc)
library(tidyverse)
library(tinytex)
library(usethis)
```

```{r}
library(tidyverse)

# Read the dataset
dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

# Count the number of establishments for each status
status_count <- dataset %>%
  group_by(`Establishment Status`) %>%
  summarize(Count = n())

# Simple plot: Number of establishments by status
ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Number of Establishments by Status",
       x = "Establishment Status", y = "Count")
```

```{r}
library(tidyverse)
library(lubridate)

dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

dataset <- dataset %>%
  mutate(InspectionYear = year(ymd(`Inspection Date`)))

total_infractions_yearly <- dataset %>%
  filter(!is.na(`Infraction Details`)) %>%
  group_by(InspectionYear) %>%
  summarize(TotalInfractions = n())

g1 <- ggplot(total_infractions_yearly, aes(x = InspectionYear, y = TotalInfractions)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Total Infractions Per Year",
       x = "Year", y = "Total Infractions")

# Load necessary libraries
library(tidyverse)
library(lubridate)

# Read the dataset
dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

# Function to calculate mode (most frequent value)
calculate_mode <- function(x) {
  ux <- unique(na.omit(x))
  ux[which.max(tabulate(match(x, ux)))]
}

# Calculating the mode of 'Min. Inspections Per Year' for each year
yearly_mode_inspections <- dataset %>%
  mutate(Year = year(ymd(`Inspection Date`))) %>%
  group_by(Year) %>%
  summarize(ModeMinInspections = calculate_mode(`Min. Inspections Per Year`))

# Plotting the graph: Most frequent 'Min. Inspections Per Year' per year
g2 <- ggplot(yearly_mode_inspections, aes(x = Year, y = ModeMinInspections)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Most Frequent 'Min. Inspections Per Year' Per Year",
       x = "Year", y = "Most Frequent Min. Inspections")

dataset <- dataset %>%
  filter(!is.na(`Min. Inspections Per Year`), !is.na(`Inspection Date`))

# Assuming the cleaned_dataset or dataset (with imputed values) is used

# Calculating the mode of 'Min. Inspections Per Year' for each year
yearly_mode_inspections <- dataset %>%
  mutate(Year = year(ymd(`Inspection Date`))) %>%
  group_by(Year) %>%
  summarize(ModeMinInspections = calculate_mode(`Min. Inspections Per Year`))

# Plotting the graph
g2 <- ggplot(yearly_mode_inspections, aes(x = Year, y = ModeMinInspections)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Most Frequent 'Min. Inspections Per Year' Per Year",
       x = "Year", y = "Most Frequent Min. Inspections")

g2

g1
g2

```

```{r}
library(tidyverse)
library(stringr)

dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

# Clean the 'Severity' column by removing the prefixes
dataset$Severity <- str_remove(dataset$Severity, "^[A-Z] - ")

# Count the number of different severities
severity_counts <- dataset %>%
  count(Severity, name = "Total_Count")

# Display the counts of different severities
severity_counts

```

```{r}
# Load necessary libraries
library(tidyverse)
library(stringr)

# Read the dataset
dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

# Clean the 'Severity' column by removing prefixes and "NA - Not Applicable"
dataset <- dataset %>%
  mutate(Severity = case_when(
    Severity == "NA - Not Applicable" ~ "Not Applicable",
    TRUE ~ str_remove(Severity, "^[A-Z]+ - ")
  ))

# Count the number of different severities, excluding "Not Applicable"
severity_counts <- dataset %>%
  filter(Severity != "Not Applicable") %>%
  count(Severity, name = "Total_Count")

# Display the counts of different severities
severity_counts

```

### 2024-01-15 Report (Codes + Texts).qmd
Rename Report.qmd to Report (Codes + Texts).qmd

---
title: "Toronto Dinesafe Open data"
subtitle: "A public health environment that is constantly improving"
author: 
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/DataAnalysis_TorontoOpenData.git."
date: January 14, 2024
date-format: long
abstract: "This report will analyze publicly available data made available to the public by Toronto City Hall. Hygiene indicators in public catering have been under serious scrutiny from the public for many years. Excellent catering hygiene conditions will drive the common development of society and economy."
format: pdf
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Toronto’s vibrant dining scene is a cornerstone of the city’s culture and economy, which is why residents and visitors require strict health screenings to ensure public safety. This report examines the Dinesafe data set provided by Toronto Public Data through the data analysis software R to conduct a comprehensive analysis of Toronto restaurant hygiene pass rates and Infractions. We further focus on the lowest inspection counts that are most frequent each year, revealing patterns that shed light on the city's public health standards.

# Data Analysis

The Dinesafe dataset is publicly available from the City of Toronto [@TorontoDineSafe2021] and covers several years, with key variables such as establishment type, inspection dates, and infraction details. Prior to analysis, we resolved missing data and extracted relevant temporal components for annual trend analysis.

```{r}
#| echo: false
#| message: false
library(tidyverse)

dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

status_count <- dataset %>%
  group_by(`Establishment Status`) %>%
  summarize(Count = n())

ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Number of Establishments by Status",
       x = "Establishment Status", y = "Count")
```

We extracted the complete data through R and classified the number of institutions by Establishment Status in Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass". Taken together the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass". "Conditional pass" indicates that a certain number of businesses must address certain issues to fully comply with health standards. This visual demonstrates that the majority of businesses inspected are adhering to the required health and safety guidelines, which is very good news for the Toronto public.

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(lubridate)

dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

dataset <- dataset %>%
  mutate(InspectionYear = year(ymd(`Inspection Date`)))

total_infractions_yearly <- dataset %>%
  filter(!is.na(`Infraction Details`)) %>%
  group_by(InspectionYear) %>%
  summarize(TotalInfractions = n())

g1 <- ggplot(total_infractions_yearly, aes(x = InspectionYear, y = TotalInfractions)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Total Infractions Per Year",
       x = "Year", y = "Total Infractions")

g1
```

Furthermore, we created a line chart by analyzing the number of Total Infractions in the data set over the years. It can be clearly seen from the chart that the total number of violations judged by catering companies each year has begun to show a downward trend after the early cycle of 2023; this shows that Toronto's public health safety is becoming excellent and more secure for people's health.

```{r}
#| echo: false
#| message: false
library(tidyverse)
library(stringr)

dataset <- read_csv("/cloud/project/data/Dinesafe.csv")

dataset <- dataset %>%
  mutate(Severity = case_when(
    Severity == "NA - Not Applicable" ~ "Not Applicable",
    TRUE ~ str_remove(Severity, "^[A-Z]+ - ")
  ))

severity_counts <- dataset %>%
  filter(Severity != "Not Applicable") %>%
  count(Severity, name = "Total_Count")

severity_counts
```

Finally, we conduct statistical analysis on the extent of these Total Infractions companies. The results show that most of the businesses with violations are at the "Minor" level (27792); this means that their violations of the hygiene assessment are minor and can be improved and reach the passing standard in a relatively faster time, which is a positive representation of Infractions The overall trend can improve faster. Despite this, data shows that 14,260 companies still face "Significant" level of sanitation infractions; this means that they may need to spend more time and cost to improve sanitation conditions to provide better protection to the public.

# Discussion and Conclusion

In summary, the purpose of this report is to examine the Dinesafe data set provided by Toronto Public Data [@TorontoDineSafe2021] through the statistical programming language R [@citeR] to conduct a comprehensive analysis of Toronto restaurant hygiene pass rates and Infractions. Include packages @AER, @future, @gitcreds, @knitr, @lintr, @renv, @reprex, @styler, @tictoc, @tidyverse, @tinytex, and @usethis.

Analysis of data sets through R and analysis of trends over time provide a valuable perspective for assessing the effectiveness of public health policies. We found that the number of infractions in Toronto restaurants is declining year over year, which means more restaurants are meeting health standards and open to the public. The highest proportion of restaurants that failed were rated "minor," indicating that more restaurants have the opportunity to improve their hygiene in a short period of time to comply with Toronto Health standards. Taken together, the above information is a very positive feedback for Toronto residents and tourists because Toronto's public restaurants are becoming healthy and standardized and increasing in number to promote economic development.

Through the analysis of the Torotno DineSafe data set, cleaning and categorizing the "severity" data allows us to conduct more granular analysis, such as in this report, making it more clear to highlight the extent of specific areas of food safety that require improvement. Distributed so that people can better analyze the hygiene status of the restaurant industry.

While this report provides a handful of solid insights, it also acknowledges the limitations of the dataset, such as potential reporting bias and missing values that may exist in the dataset. Future research could incorporate longitudinal data analysis to capture the evolution of compliance and noncompliance. In addition, this report is only a simple analysis produced for INF312 TUT2, missing many more aspects and details of description and judgment; therefore, if more accurate results are needed for the current hygiene and health of restaurants in Toronto, a more detailed and professional analysis is required. 
\newpage

# References

### 2024-01-22 Outputs Folder/paper_final.qmd
Add files via upload

---
title: "Deciphering Toronto's Food Safety in Economic Context"
subtitle: "A public health environment that is constantly improving"
author: 
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git"
date: January 21, 2024
date-format: long
abstract: "This report uses the R programming language to analyze a data set of food safety inspections at Toronto dining establishments and further uses additional data sets to correlate food safety trends with an economic indicator such as the Consumer Price Index (CPI). Our analysis shows a limited reduction in sanitation infractions; this potentially means food safety compliance is improving in the City of Toronto. Additionally, the analysis suggests some potential links between economic downturns and increases in health infractions. This report is an appropriate reference resource for policymakers, and it overall highlights the inextricable relationship between economic health and public safety."
format: pdf
toc: true
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Toronto’s vibrant dining scene is a cornerstone of the city’s culture and economy, which is why residents and visitors need strict health screenings to ensure food safety standards. Yet the public often overlooks the fact that dining establishments must adhere to strict health standards.

Over the years, public health and safety in the food service industry have received increasing attention since foodborne illness can have serious and sometimes fatal consequences. Each year, Toronto Public Health monitors, inspectcs and reports on health and safety standards at food establishments across the city; this data is fully publicly available to allow the public to have a complete understanding of the current situation. Nonetheless, among the existing studies, there are few professional analysis of Toronto restaurant food safety data, or a lack of research that directly correlates economic factors with food safety standards.

This report conducts in-depth analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto]. We use the programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details. In addition, this report extends further to examine how economic changes indicated by external factors such as the Consumer Price Index (CPI) are related to changes in public health standards within the food industry. Through analysis, we found that restaurants are more likely to violate food safety regulations, and there is a correlation between CPI fluctuations and infraction behaviours. That is, when CPI decreases, the number of infraction behaviours will increase.

Based on the above research, we aim to provide the public with further nuanced insights, inform policy decisions, and ultimately contribute to the ongoing conversation about public health and safety in urban food landscapes.

# Dataset Explanation
In this report, We used the programming language R [@citeR] and library packages include `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis] to conduct in-depth analysis from multiple aspects. Prior to analysis, we resolved missing data and extracted relevant temporal components for annual trend analysis.

## Dinesafe
The Dinesafe dataset is publicly available from the City of Toronto [@opendatatoronto] and covers several years from 2022 to 2024, with key variables such as establishment type, inspection dates, and infraction details regarding the health status of public restaurants shown in @tbl-Sample_of_DineSafe_Data.

```{r}
#| echo: false
#| label: tbl-Sample_of_DineSafe_Data
#| tbl-cap: Sample of DineSafe Data
#| message: false
#| warning: false
library(readr)
library(dplyr)
library(knitr)
library(kableExtra)

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")

sample_dataset <- dataset %>%
  select(`Establishment Status`, `Min. Inspections Per Year`, Severity) %>%
  head()

kable(sample_dataset, 
      col.names = c("Establishment Status", "Minimum Inspections Per Year", "Severity")) %>%
  kable_styling(full_width = FALSE, position = "center")

```

## Toronto's Dashboard Key Indicators
In addition, this report hopes to find the connection between public health trends and society from a broader perspective. To this end, we further searched for the dataset of Toronto's Dashboard Key Indicators through @opendatatoronto. This dataset contains many economic-related indicators that the Toronto government is open to the public; we extracted the relevant time period of the Consumer Price Index (CPI) that is consistent with dinesafe for analysis.

# Data Analysis Result

## Toronto Dine Safe Pass Status
We extracted the complete data through R and classified the number of institutions by Establishment Status in Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown. Taken together the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass". "Conditional pass" indicates that a certain number of businesses must address certain issues to fully comply with health standards. This visual demonstrates that the majority of businesses inspected are adhering to the required health and safety guidelines.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-Number_of_Establishments_by_Status
#| fig-cap: Number of Establishments by Status
library(tidyverse)

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")

status_count <- dataset %>%
  group_by(`Establishment Status`) %>%
  summarize(Count = n())

ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
       x = "Establishment Status", y = "Count")
```
\break
## Infraction Data
In addition to the companies that passed above, the DineSafe data set also shows many companies whose food safety inspections were judged to be infraction. We conduct statistical analysis on the extent of these Total Infractions companies [@tbl-total_count_severity_levels]. The results show that most of the businesses that have the amount of 28030 with violations are at the "Minor" level; this means that their violations of the hygiene assessment are minor and can be improved and reach the passing standard in a relatively faster time, which is a positive representation of Infractions The overall trend can improve faster. Despite this, data shows that 14379 companies still face a "Significant" level of sanitation infractions; this means that they may need to spend more time and cost to improve sanitation conditions to provide better protection to the public.

\break
```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-total_count_severity_levels
#| tbl-cap: Total Count of different Severity levels
library(readr)
library(dplyr)
library(knitr)
library(kableExtra)

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")

dataset <- dataset %>%
  mutate(Severity = str_remove_all(Severity, "^NA - |^[A-Z]+ - "))

severity_counts <- dataset %>%
  filter(Severity != "" & !is.na(Severity)) %>%
  count(Severity, name = "Total_Count")

kable(severity_counts, 
      col.names = c("Severity", "Total Count")) %>%
  kable_styling(full_width = FALSE, position = "center", bootstrap_options = c("striped", "hover"))

```

## Infraction Type Assessment
Since there are too many types of companies, we selected the three most numerous categories for bar chart comparison [@fig-distribution_by_severity]. They are Supermarket, Restaurant, and Food Take Out. Each category is divided into four violation severity levels: Minor, Significance, Crucial and Not Applicable (NA) further divided.

According to the analysis, we found that the category level of Minor has the highest number of infractions among all types of places, and the frequency of infractions in restaurants is particularly high. This suggests that while minor infractional behaviors are common, they are usually not serious and can be corrected with minimal intervention.

Compared to minor infractions, the number of infractions belonging to NA is relatively small. This category represents situations where standard infraction categories do not apply, perhaps due to specific exemptions or non-standard operations within the enterprise.

Significantly fewer infractions were labeled major than minor ones, which may indicate that more serious health risks are less likely to occur. Nonetheless, the presence of these infractions indicates the need for improvements to ensure public health.

The last infraction labeled Crucial is the least common, indicating relatively few cases that pose an immediate and serious health risk. However, their presence, especially in restaurants and supermarkets, highlights the importance of strict and regular inspections to prevent such significant risks to public health.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-distribution_by_severity
#| fig-cap: Distribution of Food Inspection Infractions by Severity
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)  # readr is used for read_csv

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")

selected_severity_levels <- c("C - Crucial", "M - Minor", "NA - Not Applicable", "S - Significant")
top_types <- dataset %>%
  count(`Establishment Type`) %>%
  top_n(3, n) %>%
  pull(`Establishment Type`)

dataset_filtered <- dataset %>%
  filter(Severity %in% selected_severity_levels, `Establishment Type` %in% top_types)

ggplot(dataset_filtered, aes(x = `Establishment Type`, fill = Severity)) +
  geom_bar(alpha = 0.8, position = position_dodge(width = 0.75)) +
  theme_minimal() +
  facet_wrap(~Severity, ncol = 1, scales = "free_y") +
  coord_flip() +
  labs(x = "Establishment Type", y = "Count of Infractions")

```
\break
## Actions from Infraction Companies
To further evaluate the actions taken by these violating companies after being flagged, we used the “Minor” level with the most violations for analysis and created a pie chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions]. The chart is dominated by the blue proportion, which represents the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action consider by Toronto Public Health; instead, businesses are given the opportunity to make amends within a certain period.

```{r}
#| echo: false
#| message: false
#| label: fig-distribution_minor_severity_infractions
#| fig-cap: Distribution of Food Inspection Infractions by Severity
#| warning: false
library(ggplot2)
library(dplyr)
library(opendatatoronto)
library(lubridate)

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")

minor_severity_actions <- dataset %>%
  filter(Severity == "M - Minor") %>%
  count(Action) %>%
  arrange(desc(n))

ggplot(minor_severity_actions, aes(x = "", y = n, fill = Action)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") + # This creates a pie chart, remove this line for a stacked bar chart
  theme_minimal() +
  labs(
       x = "Action", y = "Count")

```
\break
## Potential Relationship between CPI and Number of Infractions
CPI is an economic indicator that measures changes in the inflation rate or cost of living in a country or region [@goldberg2010sensitivity]. The relevant data for this CPI comes from Toronto's Dashboard Key Indicators data set published by the City of Toronto @opendatatoronto. We generated two line charts [@fig-trend_comparison_of_cpi_and_infractions], showing the CPI trend and the Infraction number trend respectively. Observation shows that when CPI decreases, corporate hygiene violations will increase. Although there may not be a direct causal relationship between the two, their trends may be affected by similar external factors. For example, a decline in CPI is usually a sign of a recession, which indicates that Toronto residents will potentially have more difficulties financially; therefore, businesses will spend less money on maintaining food safety conditions in order to stabilize the economic situation.

```{r}
#| echo: false
#| message: false
#| eval: true
#| label: fig-trend_comparison_of_cpi_and_infractions
#| fig-cap: "Trend comparison of CPI and Infractions"
#| fig-subcap: ["CPI Trend","Infractions Trend"]
#| layout-ncol: 2
#| warning: false
library(ggplot2)
library(dplyr)
library(readr)

cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal2.csv") %>%
  filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
  rename(CPI = measure_value) %>%
  mutate(Year = as.numeric(year)) %>%
  group_by(Year) %>%
  summarize(Average_CPI = mean(CPI, na.rm = TRUE))

dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv") %>%
  mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
  mutate(Year = as.numeric(format(Inspection_Date, "%Y"))) %>%
  group_by(Year) %>%
  summarize(Total_Infractions = n())

common_years <- intersect(cpi_data$Year, dataset$Year)

cpi_common_years <- filter(cpi_data, Year %in% common_years)
dinesafe_common_years <- filter(dataset, Year %in% common_years)

cpi_plot <- ggplot(cpi_common_years, aes(x = Year, y = Average_CPI)) +
  geom_line(color = "red") +
  scale_y_continuous(name = "CPI (% Change)", labels = scales::percent) +
  labs() +
  theme_minimal()

dinesafe_plot <- ggplot(dinesafe_common_years, aes(x = Year, y = Total_Infractions)) +
  geom_line(color = "blue") +
  scale_y_continuous(name = "Total Infractions") +
  labs() +
  theme_minimal()

print(cpi_plot)
print(dinesafe_plot)

```
\break
# Discussion and Conclusion
This report examines the state of public food safety in Toronto using two different data sets provided by Toronto City Hall: Dinesafe and Toronto's Dashboard Key Indicators through the statistical programming language R [@citeR] to conduct a comprehensive analysis of Toronto restaurant hygiene pass rates and Infractions. Include packages `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis]. We focused on an in-depth study of the DineSafe data and concluded that the number of violations showed a clear downward trend within the constraints of the dataset. It is worth noting that the data shows that most enterprises have passed or conditionally passed the assessment of Toronto Public Health. The vast majority of enterprises marked as infractions are at the "Minor" level. They are usually notified of rectification without further punish.

The high compliance rates above may reflect the effectiveness of Toronto Public Health's strategy; however, persistence levels are labelled "Significance" and "Crucial" in the data. Although the number of severe levels of these health conditions is small, relevant administrative departments and even the public must continue to be vigilant about the need and take appropriate measures to contain it.

Beyond this, our comparative analysis of DineSafe and Toronto's Dashboard Key Indicators containing the CPI data set further illuminates the interplay between economic indicators and public health outcomes. A potential economic downturn reflected in the fall in the Consumer Price Index could lead to tighter budgets within food businesses, potentially compromising hygiene standards, according to analysis from a data visualization made in R.

Despite some valid findings, we acknowledge the limitations inherent in this report and the two datasets it used. First, we could not examine these data sets provided by the City of Toronto for potential biases and incomplete content that could distort interpretations of actual trends. Secondly, CPI is a complex indicator that may be affected by various economic activities and policy changes. These factors have not been fully considered in this report. Therefore, future research efforts should use broader data sets and more precise analyses to provide a more convincing picture of public food security in Toronto.
\newpage

# References

### 2024-01-22 Scripts Folder/scripts/02-data_cleaning.R

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Clean Dinesafe and CPI datas from OpenDataToronto
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read.csv("/cloud/project/Inputs Folder/Data/raw_data.csv")
+
+cleaned_data <- dataset %>%
+  filter(!is.na(`Establishment Name`)) %>%
+  mutate(new_column = tolower(`Establishment Name`))
+
+#### Save data ####
+write.csv(dataset, "analysis_data_final1X.csv")
+
+#### Clean data ####
+raw_data2 <- read.csv("/cloud/project/Inputs Folder/Data/raw_data2.csv")
+
+cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
+
+cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
+
+#### Save data ####
+write.csv(cpi_data, "analysis_data_final2.csv")

### 2024-01-22 Scripts Folder/scripts/03-test_data.R

@@ -1,0 +1,78 @@
+#### Preamble ####
+# Purpose: Tests whether dataset can operate normally
+# Author: Tianen (Evan) Hao
+# Date: 15 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+
+#### Test dinesafe data ####
+library(tidyverse)
+
+# Use the simulated DineSafe data set
+set.seed(123)
+simulated_dataset <- tibble(
+  `Establishment ID` = 1:100,
+  `Establishment Name` = paste("Establishment", 1:100),
+  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
+  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
+  `Infractions` = sample(0:5, 100, replace = TRUE)
+)
+
+status_count <- simulated_dataset %>%
+  group_by(`Establishment Status`) %>%
+  summarize(Count = n())
+
+ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  theme_minimal() +
+  labs(title = "Number of Establishments by Status",
+       x = "Establishment Status", y = "Count")
+
+library(dplyr)
+
+inspection_summary <- simulated_dataset %>%
+  group_by(`Establishment Status`) %>%
+  summarise(Count = n()) %>%
+  mutate(Percentage = (Count / sum(Count)) * 100)
+
+print(inspection_summary)
+
+
+#### Test Toronto's Dashboard Key Indicators data set####
+
+library(dplyr)
+library(tibble)
+
+# Use the simulated Toronto's Dashboard Key Indicators data set
+set.seed(123)
+
+cpi_data <- tibble(
+  Year = 2010:2023, 
+  Average_CPI = runif(length(Year), min = 0, max = 5) 
+)
+
+print(cpi_data)
+
+library(ggplot2)
+
+ggplot(cpi_data, aes(x = Year, y = Average_CPI)) +
+  geom_line() +
+  labs(title = "Annual CPI Trend", x = "Year", y = "Average CPI")
+
+cpi_data <- cpi_data %>%
+  mutate(Growth_Rate = (Average_CPI - lag(Average_CPI)) / lag(Average_CPI))
+
+print(cpi_data)
+
[…]

### 2024-01-22 Outputs Folder 2/paper_final.qmd
Add files via upload

--- 2024-01-22 Outputs Folder/paper_final.qmd
@@
 ---
-title: "Deciphering Toronto's Food Safety in Economic Context"
+title: "Analyzing Toronto's Dining Scene: Public Food Safety and Economic Indicators in Focus"
 subtitle: "A public health environment that is constantly improving"

### 2024-01-22 Outputs Folder/paper_final.qmd
Fixed some grammatical issues in Paper

@@
 # Dataset Explanation
-In this report, We used the programming language R [@citeR] and library packages include `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis] to conduct in-depth analysis from multiple aspects.
+In this report, we used the programming language R [@citeR] and library packages include `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis] to conduct in-depth analysis from multiple aspects.
 Prior to analysis, we resolved missing data and extracted relevant temporal components for annual trend analysis.
@@
 CPI is an economic indicator that measures changes in the inflation rate or cost of living in a country or region [@goldberg2010sensitivity].
+In the early days of writing this report, the AI language model ChatGPT 4 [@openai2023chatgpt4] provided the author with inspiration for relevant economic indicators; the author finally selected CPI as a suitable indicator.
 The relevant data for this CPI comes from Toronto's Dashboard Key Indicators data set published by the City of Toronto @opendatatoronto.
-We generated two line charts [@fig-trend_comparison_of_cpi_and_infractions], showing the CPI trend and the Infraction number trend respectively.
+We generated two line charts [@fig-trend_comparison_of_cpi_and_infractions], showing the CPI trend and the infraction number trend respectively.
 Observation shows that when CPI decreases, corporate hygiene violations will increase.

### 2024-01-23 Outputs Folder/paper_final.qmd
Fixed typos and update the appendix

--- 2024-01-22 Outputs Folder/paper_final.qmd
@@
 ---
-title: "Deciphering Toronto's Food Safety in Economic Context"
-subtitle: "A public health environment that is constantly improving"
+title: "Analyzing Toronto's Dining Scene: Public Food Safety and Economic Indicators in Focus"
 author: 
@@
 date-format: long
-abstract: "This report uses the R programming language to analyze a data set of food safety inspections at Toronto dining establishments and further uses additional data sets to correlate food safety trends with an economic indicator such as the Consumer Price Index (CPI).
-Our analysis shows a limited reduction in sanitation infractions; this potentially means food safety compliance is improving in the City of Toronto.
-Additionally, the analysis suggests some potential links between economic downturns and increases in health infractions.
-This report is an appropriate reference resource for policymakers, and it overall highlights the inextricable relationship between economic health and public safety."
+abstract: "Nowadays, as the market economy continues to thrive, people have more diverse choices of edible products and are also more aware of their potential health risks.
+This report uses the open-source statistical programming language R to analyze a data set of food safety inspections at Toronto dining establishments and further uses additional data sets to correlate food safety trends with an economic indicator such as the Consumer Price Index (CPI).
+The analysis shows a limited reduction in sanitation infractions; this potentially means food safety compliance is improving in the City of Toronto.
+Additionally, the analysis suggests some potential links between economic downturns and increases in health infractions."
 format: pdf
@@
 
-Toronto’s vibrant dining scene is a cornerstone of the city’s culture and economy, which is why residents and visitors need strict health screenings to ensure food safety standards.
-Yet the public often overlooks the fact that dining establishments must adhere to strict health standards.
+Toronto’s vibrant dining scene is a cornerstone of the city’s culture and economy; therefore, residents and visitors need strict health screenings to ensure food safety standards.
+Nevertheless, the public often overlooks that dining establishments must adhere to strict health standards.
 
-Over the years, public health and safety in the food service industry have received increasing attention since foodborne illness can have serious and sometimes fatal consequences.
-Each year, Toronto Public Health monitors, inspectcs and reports on health and safety standards at food establishments across the city; this data is fully publicly available to allow the public to have a complete understanding of the current situation.
+Over the years, public health and safety in the food service industry have received increasing attention since foodborne illness can have severe and sometimes fatal consequences.
+Each year, Toronto Public Health monitors, inspects, and reports on health and safety standards at food establishments across the city; this data is fully publicly available to allow the public to understand the current situation.
 Nonetheless, among the existing studies, there are few professional analysis of Toronto restaurant food safety data, or a lack of research that directly correlates economic factors with food safety standards.
 
-This report conducts in-depth analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto].
-We use the programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details.
-In addition, this report extends further to examine how economic changes indicated by external factors such as the Consumer Price Index (CPI) are related to changes in public health standards within the food industry.
-Through analysis, we found that restaurants are more likely to violate food safety regulations, and there is a correlation between CPI fluctuations and infraction behaviours.
-That is, when CPI decreases, the number of infraction behaviours will increase.
+This report conducts an in-depth analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto].
+The author use the open-source statistical programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details.
+In addition, this report further examines how economic changes indicated by external factors such as the Consumer Price Index (CPI) are related to changes in public health standards within the food industry.
+Through analysis, results found that restaurants are more likely to violate food safety regulations, and there is a correlation between CPI fluctuations and infraction behaviours.
+The trend shows that the number of infraction behaviours will increase when CPI decreases.
 
-Based on the above research, we aim to provide the public with further nuanced insights, inform policy decisions, and ultimately contribute to the ongoing conversation about public health and safety in urban food landscapes.
+Based on the above research, the report aim to provide the public with further nuanced insights, inform policy decisions, and ultimately contribute to the ongoing conversation about public health and safety in urban food landscapes.
 
 # Dataset Explanation
-In this report, We used the programming language R [@citeR] and library packages include `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis] to conduct in-depth analysis from multiple aspects.
-Prior to analysis, we resolved missing data and extracted relevant temporal components for annual trend analysis.
+In this report, the author used the open-source statistical programming language R [@citeR] and library packages include `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis] to conduct in-depth analysis from multiple aspects.
 
 ## Dinesafe
-The Dinesafe dataset is publicly available from the City of Toronto [@opendatatoronto] and covers several years from 2022 to 2024, with key variables such as establishment type, inspection dates, and infraction details regarding the health status of public restaurants shown in @tbl-Sample_of_DineSafe_Data.
-
+The Dinesafe dataset is publicly available from the City of Toronto [@opendatatoronto] and has many related variables.
+In this report, some key variables are mainly selected, including inspection dates, establishment type, establishment status, severity, action, and minimum inspections per year, as @tbl-Sample_of_DineSafe_Data shown.
+Inspection dates represent the time range for data set inspection; this report mainly selects the range from January 2022 to January 2024 for spot inspection analysis.
+Establishment type represents the relevant business type when Toronto Public Health inspects, such as a restaurant, food store or food takeout.
+Establishment status mainly reflects whether an enterprise has passed the food safety inspection.
+In this data set, only relevant enterprises that successfully passed or passed conditionally are displayed.
+Although it does not directly reflect the situation of failed companies, the severity variable in the data set shows that these companies that passed also have infraction problems in different situations, and the action variable records the different actions taken by these companies after infractions.
+Finally, minimum inspections per year serve as a reference for analyzing food safety status and corporate behaviour.
+Considering the hugeness of the dataset data, please refer to \ref{app:clean-dinesafe-dataset}.
 ```{r}
@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")
+dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
 
@@
 ```
-
 ## Toronto's Dashboard Key Indicators
 In addition, this report hopes to find the connection between public health trends and society from a broader perspective.
-To this end, we further searched for the dataset of Toronto's Dashboard Key Indicators through @opendatatoronto.
-This dataset contains many economic-related indicators that the Toronto government is open to the public; we extracted the relevant time period of the Consumer Price Index (CPI) that is consistent with dinesafe for analysis.
+To this end, the report further searched for Toronto's Dashboard Key Indicators dataset through @opendatatoronto.
+This dataset contains many economic-related indicators that the Toronto government is open to the public; the author selected the relevant period between January 2022 and January 2023 of the Consumer Price Index (CPI), including numerical values and years that are consistent with Dinesafe for analysis.
+Considering the hugeness of the dataset data, please refer to \ref{app:clean-cpi-dataset}.
 
@@
 ## Toronto Dine Safe Pass Status
-We extracted the complete data through R and classified the number of institutions by Establishment Status in Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown.
-Taken together the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass".
-"Conditional pass" indicates that a certain number of businesses must address certain issues to fully comply with health standards.
-This visual demonstrates that the majority of businesses inspected are adhering to the required health and safety guidelines.
+The report extracted the complete data through R and classified the number of institutions by Establishment Status in the Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown.
+Taken together, the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards.
+This visual demonstrates that most businesses inspected adhere to the required health and safety guidelines.
 
@@
 ```
-\break
+
 ## Infraction Data
-In addition to the companies that passed above, the DineSafe data set also shows many companies whose food safety inspections were judged to be infraction.
-We conduct statistical analysis on the extent of these Total Infractions companies [@tbl-total_count_severity_levels].
+
+In one of the companies that passed above, the DineSafe data set shows many companies whose food safety inspections were judged to be infractions.
+The report analyzes the extent of these total infractions companies [@tbl-total_count_severity_levels].
 The results show that most of the businesses that have the amount of 28030 with violations are at the "Minor" level; this means that their violations of the hygiene assessment are minor and can be improved and reach the passing standard in a relatively faster time, which is a positive representation of Infractions The overall trend can improve faster.
-Despite this, data shows that 14379 companies still face a "Significant" level of sanitation infractions; this means that they may need to spend more time and cost to improve sanitation conditions to provide better protection to the public.
+Despite this, data shows that 14379 companies still face a "Significant" level of sanitation infractions; this means they may need to spend more time and money to improve sanitation conditions to provide better protection to the public.
 
-\break
 ```{r}
@@
 ## Infraction Type Assessment
-Since there are too many types of companies, we selected the three most numerous categories for bar chart comparison [@fig-distribution_by_severity].
+Since there are too many types of companies, the report selected the three most numerous categories for bar chart comparison [@fig-distribution_by_severity].
 They are Supermarket, Restaurant, and Food Take Out.
-Each category is divided into four violation severity levels: Minor, Significance, Crucial and Not Applicable (NA) further divided.
+Each category is divided into four violation severity levels: Minor, Significance, Crucial and Not Applicable (NA), which are further divided.
+This report referred to the relevant codes of the @rstats_top50ggplot2 website when writing the code for this bar chart.
 
-According to the analysis, we found that the category level of Minor has the highest number of infractions among all types of places, and the frequency of infractions in restaurants is particularly high.
-This suggests that while minor infractional behaviors are common, they are usually not serious and can be corrected with minimal intervention.
+The analysis found that the category level of Minor has the highest number of infractions among all types of places, and the frequency of infractions in restaurants is exceptionally high.
+This suggests that while minor infractional behaviours are common, they are usually not serious and can be corrected with minimal intervention.
 
-Compared to minor infractions, the number of infractions belonging to NA is relatively small.
+Compared to minor infractions, the number of infractions belonging to NA is relatively tiny.
 This category represents situations where standard infraction categories do not apply, perhaps due to specific exemptions or non-standard operations within the enterprise.
 
-Significantly fewer infractions were labeled major than minor ones, which may indicate that more serious health risks are less likely to occur.
+Significantly fewer infractions were labelled major than minor ones, which may indicate that more serious health risks are less likely to occur.
 Nonetheless, the presence of these infractions indicates the need for improvements to ensure public health.
 
-The last infraction labeled Crucial is the least common, indicating relatively few cases that pose an immediate and serious health risk.
+The last infraction, Crucial, is the least common, indicating relatively few cases that pose an immediate and serious health risk.
 However, their presence, especially in restaurants and supermarkets, highlights the importance of strict and regular inspections to prevent such significant risks to public health.
@@
 ## Actions from Infraction Companies
-To further evaluate the actions taken by these violating companies after being flagged, we used the “Minor” level with the most violations for analysis and created a pie chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
-The chart is dominated by the blue proportion, which represents the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action consider by Toronto Public Health; instead, businesses are given the opportunity to make amends within a certain period.
+To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a pie chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
+Please note that the coding of this chart refers to the "Pie Chart customization" website [@rggpiechart].
+The chart is dominated by the green proportion, representing the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action Toronto Public Health considers; instead, businesses are allowed to make amends within a certain period.
 
@@
 library(dplyr)
-library(opendatatoronto)
 library(lubridate)
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv")
+dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
 
@@
 ## Potential Relationship between CPI and Number of Infractions
+Finally, in order to observe food safety issues from a broader social perspective, the report chose to use the Consumer Price Index (CPI) for further comparison.
 CPI is an economic indicator that measures changes in the inflation rate or cost of living in a country or region [@goldberg2010sensitivity].
+In the early stages of writing this report, the AI language model ChatGPT 4 [@openai2023chatgpt4] provided the author with inspiration for relevant economic indicators; the author finally selected CPI as a suitable indicator.
 The relevant data for this CPI comes from Toronto's Dashboard Key Indicators data set published by the City of Toronto @opendatatoronto.
-We generated two line charts [@fig-trend_comparison_of_cpi_and_infractions], showing the CPI trend and the Infraction number trend respectively.
+The report generated two line charts [@fig-trend_comparison_of_cpi_and_infractions], showing the CPI trend and the infraction number trend, respectively.
 Observation shows that when CPI decreases, corporate hygiene violations will increase.
 Although there may not be a direct causal relationship between the two, their trends may be affected by similar external factors.
-For example, a decline in CPI is usually a sign of a recession, which indicates that Toronto residents will potentially have more difficulties financially; therefore, businesses will spend less money on maintaining food safety conditions in order to stabilize the economic situation.
+For example, a decline in CPI is usually a sign of a recession [@goldberg2010sensitivity], indicating that Toronto residents will potentially have more financial difficulties; therefore, businesses will spend less money on maintaining food safety conditions to stabilize the economic situation.
 
@@
 
-cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal2.csv") %>%
+cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv") %>%
   filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/AnalysisDataFinal1.csv") %>%
+dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
   mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
@@
 This report examines the state of public food safety in Toronto using two different data sets provided by Toronto City Hall: Dinesafe and Toronto's Dashboard Key Indicators through the statistical programming language R [@citeR] to conduct a comprehensive analysis of Toronto restaurant hygiene pass rates and Infractions.
-Include packages `AER`[@AER], `future`[@future], `gircreds`[@gitcreds], `knitr`[@knitr], `lintr`[@lintr], `renv`[@renv], `reprex`[@reprex], `styler`[@styler], `tictoc`[@tictoc], `tidyverse`[@tidyverse], `tinytex`[@tinytex], and `usethis`[@usethis].
-We focused on an in-depth study of the DineSafe data and concluded that the number of violations showed a clear downward trend within the constraints of the dataset.
+The analysis focused on an in-depth study of the DineSafe data and concluded that the number of infractions showed a clear downward trend within the constraints of the dataset.
 It is worth noting that the data shows that most enterprises have passed or conditionally passed the assessment of Toronto Public Health.
-The vast majority of enterprises marked as infractions are at the "Minor" level.
-They are usually notified of rectification without further punish.
+The vast majority of enterprises marked as infractions are at the "Minor" level; they are usually notified of rectification without further punishment.
 
-The high compliance rates above may reflect the effectiveness of Toronto Public Health's strategy; however, persistence levels are labelled "Significance" and "Crucial" in the data.
-Although the number of severe levels of these health conditions is small, relevant administrative departments and even the public must continue to be vigilant about the need and take appropriate measures to contain it.
+The high compliance rates above may reflect the effectiveness of Toronto Public Health's strategy; however, there are still some companies labelled as "Significance" and "Crucial" in the data.
+Although the number of severe levels of these food safety conditions is small, relevant administrative departments and the public must continue to be vigilant about the need and take appropriate measures to contain it.
 
@@
 
-Despite some valid findings, we acknowledge the limitations inherent in this report and the two datasets it used.
-First, we could not examine these data sets provided by the City of Toronto for potential biases and incomplete content that could distort interpretations of actual trends.
+Despite some valid findings, the author acknowledges the limitations inherent in this report and the two datasets it used.
+First, the report could not examine these data sets provided by the City of Toronto for potential biases and incomplete content that could distort interpretations of actual trends.
 Secondly, CPI is a complex indicator that may be affected by various economic activities and policy changes.
 These factors have not been fully considered in this report.
-Therefore, future research efforts should use broader data sets and more precise analyses to provide a more convincing picture of public food security in Toronto.
+Therefore, future research efforts should use broader data sets and more precise analyses to provide a more convincing picture of public food safety in Toronto.
+
+\appendix
+
+# Appendix {-}
+
+\section{Clean Dinesafe Dataset} 
+\label{app:clean-dinesafe-dataset} 
+Considering the huge content of the data set, this appendix contains a preview of the cleaned data set that has been used for the research.
+Original data is avaliable at https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+library(dplyr)
+library(magrittr)  
+library(kableExtra)
+library(readr)
+library(tidyr)
+library(knitr)
+
+file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"
+dataset <- read_csv(file_path)
+
+processed_dataset <- dplyr::select(dataset, 
+                                   Inspection_Dates = `Inspection Date`, 
+                                   Establishment_Type = `Establishment Type`,
+                                   Establishment_Status = `Establishment Status`,
+                                   Severity,
+                                   Action,
+                                   Min_Inspections_Per_Year = `Min. Inspections Per Year`)
+
+summary_table <- dplyr::group_by(processed_dataset, Establishment_Type, Establishment_Status) %>%
+  dplyr::summarise(Total_Count = n(), 
+                   Avg_Min_Inspections = mean(Min_Inspections_Per_Year, na.rm = TRUE)) %>%
+  dplyr::arrange(Establishment_Type, Establishment_Status)
+
+#knitr::kable(summary_table)
+
+kable(head(summary_table,20), format = "latex", longtable = TRUE, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("striped", "repeat_header"))
+
+```
+\clearpage
+\section{Clean CPI Dataset} 
+\label{app:clean-cpi-dataset} 
+Considering the huge content of the data set, this appendix contains a preview of the cleaned data set that has been used for the research.
+Original data is avaliable at https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+file_path <- "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv"
+
+cpi_data <- read_csv(file_path)
+
+cpi_filtered <- cpi_data %>%
+  filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
+  select(Year = year, CPI_Value = measure_value)
+
+kable(head(cpi_filtered,20), format = "latex", booktabs = TRUE) %>%
+  kable_styling(latex_options = "striped", position = "center")
+
+```
+
 \newpage
-
 # References

### 2024-01-23 Scripts Folder/scripts/01-download_data.R
Delete Scripts Folder/scripts directory

@@ -1,49 +1,0 @@
-#### Preamble ####
-# Purpose: Download and read Dinesafe and CPI datas from OpenDataToronto
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-library(opendatatoronto)
-
-#### Download data ####
-# get package
-package <- show_package("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
-package
-
-# get all resources for this package
-resources <- list_package_resources("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-dataset <- filter(datastore_resources, row_number()==1) %>% get_resource()
-
-#### Save data ####
-write.csv(dataset, "raw_data.csv") 
-
-#### Download data 2####
-
-# Get package
-package <- show_package("c6d64e9b-f85a-4084-be14-60cf18509203")
-# It's a good practice to check the package details, but you can also comment it out if not needed
-# package
-
-# Get all resources for this package
-resources <- list_package_resources("c6d64e9b-f85a-4084-be14-60cf18509203")
-
-# Identify datastore resources; Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# Load the first datastore resource as a sample
-# Using `slice()` instead of `filter()` to select the first resource
-data <- slice(datastore_resources, 1) %>% get_resource()
-
-# View the first few rows of the dataset
-head(data)
-         
-#### Save data 2####
-write.csv(cpi_data, "raw_data2.csv") 

### 2024-01-23 Scripts Folder/scripts/02-data_cleaning.R

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: Clean Dinesafe and CPI datas from OpenDataToronto
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean data ####
-raw_data <- read.csv("/cloud/project/Inputs Folder/Data/raw_data.csv")
-
-cleaned_data <- dataset %>%
-  filter(!is.na(`Establishment Name`)) %>%
-  mutate(new_column = tolower(`Establishment Name`))
-
-#### Save data ####
-write.csv(dataset, "analysis_data_final1X.csv")
-
-#### Clean data ####
-raw_data2 <- read.csv("/cloud/project/Inputs Folder/Data/raw_data2.csv")
-
-cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
-
-cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
-
-#### Save data ####
-write.csv(cpi_data, "analysis_data_final2.csv")

### 2024-01-23 Scripts Folder/scripts/03-test_data.R

@@ -1,78 +1,0 @@
-#### Preamble ####
-# Purpose: Tests whether dataset can operate normally
-# Author: Tianen (Evan) Hao
-# Date: 15 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-
-#### Test dinesafe data ####
-library(tidyverse)
-
-# Use the simulated DineSafe data set
-set.seed(123)
-simulated_dataset <- tibble(
-  `Establishment ID` = 1:100,
-  `Establishment Name` = paste("Establishment", 1:100),
-  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
-  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
-  `Infractions` = sample(0:5, 100, replace = TRUE)
-)
-
-status_count <- simulated_dataset %>%
-  group_by(`Establishment Status`) %>%
-  summarize(Count = n())
-
-ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
-  theme_minimal() +
-  labs(title = "Number of Establishments by Status",
-       x = "Establishment Status", y = "Count")
-
-library(dplyr)
-
-inspection_summary <- simulated_dataset %>%
-  group_by(`Establishment Status`) %>%
-  summarise(Count = n()) %>%
-  mutate(Percentage = (Count / sum(Count)) * 100)
-
-print(inspection_summary)
-
-
-#### Test Toronto's Dashboard Key Indicators data set####
-
-library(dplyr)
-library(tibble)
-
-# Use the simulated Toronto's Dashboard Key Indicators data set
-set.seed(123)
-
-cpi_data <- tibble(
-  Year = 2010:2023, 
-  Average_CPI = runif(length(Year), min = 0, max = 5) 
-)
-
-print(cpi_data)
-
-library(ggplot2)
-
-ggplot(cpi_data, aes(x = Year, y = Average_CPI)) +
-  geom_line() +
-  labs(title = "Annual CPI Trend", x = "Year", y = "Average CPI")
-
-cpi_data <- cpi_data %>%
-  mutate(Growth_Rate = (Average_CPI - lag(Average_CPI)) / lag(Average_CPI))
-
-print(cpi_data)
-
[…]

### 2024-01-23 Scripts Folder/scripts/01-download_data.R
Update and fix boolean testing

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Download and read Dinesafe and CPI datas from OpenDataToronto
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+library(opendatatoronto)
+
+#### Download data ####
+# get package
+package <- show_package("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
+package
+
+# get all resources for this package
+resources <- list_package_resources("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+dataset <- filter(datastore_resources, row_number()==1) %>% get_resource()
+
+#### Save data ####
+write.csv(dataset, "raw_data.csv") 
+
+#### Download data 2####
+
+# Get package
+package <- show_package("c6d64e9b-f85a-4084-be14-60cf18509203")
+# It's a good practice to check the package details, but you can also comment it out if not needed
+# package
+
+# Get all resources for this package
+resources <- list_package_resources("c6d64e9b-f85a-4084-be14-60cf18509203")
+
+# Identify datastore resources; Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# Load the first datastore resource as a sample
+# Using `slice()` instead of `filter()` to select the first resource
+data <- slice(datastore_resources, 1) %>% get_resource()
+
+# View the first few rows of the dataset
+head(data)
+         
+#### Save data 2####
+write.csv(cpi_data, "raw_data2.csv") 

### 2024-01-23 Scripts Folder/scripts/02-data_cleaning.R

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Clean Dinesafe and CPI datas from OpenDataToronto
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read.csv("/cloud/project/Inputs Folder/Data/raw_data.csv")
+
+cleaned_data <- dataset %>%
+  filter(!is.na(`Establishment Name`)) %>%
+  mutate(new_column = tolower(`Establishment Name`))
+
+#### Save data ####
+write.csv(dataset, "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+
+#### Clean data ####
+raw_data2 <- read.csv("/cloud/project/Inputs Folder/Data/raw_data2.csv")
+
+cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
+
+cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
+
+#### Save data ####
+write.csv(cpi_data, "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv")

### 2024-01-23 Scripts Folder/scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Clean Dinesafe and CPI datas from OpenDataToronto
+# Purpose: Clean Dinesafe and CPI dataset
 # Author: Tianen (Evan) Hao
 # Date: 14 January 2024
 # Contact: evan4830@gmail.com

### 2024-01-23 Scripts Folder/scripts/01-download_data.R
Update 01-download_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Download and read Dinesafe and CPI datas from OpenDataToronto
+# Purpose: Download and read Dinesafe and CPI data from OpenDataToronto
 # Author: Tianen (Evan) Hao
 # Date: 14 January 2024
 # Contact: evan4830@gmail.com

### 2024-01-23 Scripts Folder/scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -10,21 +10,21 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read.csv("/cloud/project/Inputs Folder/Data/raw_data.csv")
+raw_data <- read.csv("Inputs Folder/Data/raw_data.csv")
 
 cleaned_data <- dataset %>%
   filter(!is.na(`Establishment Name`)) %>%
   mutate(new_column = tolower(`Establishment Name`))
 
 #### Save data ####
-write.csv(dataset, "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+write.csv(dataset, "Inputs Folder/Data/Clean_Dinesafe_Data.csv")
 
 #### Clean data ####
-raw_data2 <- read.csv("/cloud/project/Inputs Folder/Data/raw_data2.csv")
+raw_data2 <- read.csv("Inputs Folder/Data/raw_data2.csv")
 
 cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
 
 cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
 
 #### Save data ####
-write.csv(cpi_data, "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv")
+write.csv(cpi_data, "Inputs Folder/Data/Clean_CPI_Data.csv")

### 2024-01-23 Outputs Folder/paper_final.qmd
Update paper_final.qmd

@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv")
 
@@
 
-cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv") %>%
+cpi_data <- read_csv("Inputs Folder/Data/Clean_CPI_Data.csv") %>%
   filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
+dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
   mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
@@
 
-file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"
+file_path <- "Inputs Folder/Data/Clean_Dinesafe_Data.csv"
 dataset <- read_csv(file_path)
@@
 
-file_path <- "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv"
+file_path <- "Inputs Folder/Data/Clean_CPI_Data.csv"
 

### 2024-01-23 Scripts Folder/scripts/01-download_data.R
Delete Scripts Folder/scripts directory

@@ -1,49 +1,0 @@
-#### Preamble ####
-# Purpose: Download and read Dinesafe and CPI data from OpenDataToronto
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-library(opendatatoronto)
-
-#### Download data ####
-# get package
-package <- show_package("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
-package
-
-# get all resources for this package
-resources <- list_package_resources("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-dataset <- filter(datastore_resources, row_number()==1) %>% get_resource()
-
-#### Save data ####
-write.csv(dataset, "raw_data.csv") 
-
-#### Download data 2####
-
-# Get package
-package <- show_package("c6d64e9b-f85a-4084-be14-60cf18509203")
-# It's a good practice to check the package details, but you can also comment it out if not needed
-# package
-
-# Get all resources for this package
-resources <- list_package_resources("c6d64e9b-f85a-4084-be14-60cf18509203")
-
-# Identify datastore resources; Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# Load the first datastore resource as a sample
-# Using `slice()` instead of `filter()` to select the first resource
-data <- slice(datastore_resources, 1) %>% get_resource()
-
-# View the first few rows of the dataset
-head(data)
-         
-#### Save data 2####
-write.csv(cpi_data, "raw_data2.csv") 

### 2024-01-23 Scripts Folder/scripts/03-test_data.R

@@ -1,106 +1,0 @@
-#### Preamble ####
-# Purpose: Tests whether dataset can operate normally
-# Author: Tianen (Evan) Hao
-# Date: 15 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-
-#### Test dinesafe data ####
-library(tidyverse)
-
-# Use the simulated DineSafe data set
-set.seed(123)
-simulated_dataset <- tibble(
-  `Establishment ID` = 1:100,
-  `Establishment Name` = paste("Establishment", 1:100),
-  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
-  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
-  `Infractions` = sample(0:5, 100, replace = TRUE)
-)
-
-# Test if 'Establishment Status' contains specific categories
-all(c("Pass", "Conditional Pass", "Fail") %in% simulated_dataset$`Establishment Status`)
-
-# Test the range of dates in 'Inspection Date'
-simulated_dataset$`Inspection Date` |> min() >= as.Date('2021-01-01')
-simulated_dataset$`Inspection Date` |> max() <= as.Date('2021-12-31')
-
-# Test the range of values in 'Infractions'
-simulated_dataset$Infractions |> min() >= 0
-simulated_dataset$Infractions |> max() <= 5
-
-# Test the class of 'Establishment ID'
-simulated_dataset$`Establishment ID` |> class() == "integer"
-
-# Some extra test
-status_count <- simulated_dataset %>%
-  group_by(`Establishment Status`) %>%
-  summarize(Count = n())
-
-ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
-  theme_minimal() +
-  labs(title = "Number of Establishments by Status",
-       x = "Establishment Status", y = "Count")
-
-library(dplyr)
-
-inspection_summary <- simulated_dataset %>%
-  group_by(`Establishment Status`) %>%
-  summarise(Count = n()) %>%
-  mutate(Percentage = (Count / sum(Count)) * 100)
-
-print(inspection_summary)
-
-
-#### Test Toronto's Dashboard Key Indicators data set####
-
[…]

### 2024-01-23 Scripts Folder/scripts/02-data_cleaning.R
Updated Boolean testing

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Clean Dinesafe and CPI dataset
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read.csv("Inputs Folder/Data/raw_data.csv")
+
+cleaned_data <- dataset %>%
+  filter(!is.na(`Establishment Name`)) %>%
+  mutate(new_column = tolower(`Establishment Name`))
+
+#### Save data ####
+write.csv(dataset, "Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+
+#### Clean data ####
+raw_data2 <- read.csv("Inputs Folder/Data/raw_data2.csv")
+
+cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
+
+cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
+
+#### Save data ####
+write.csv(cpi_data, "Inputs Folder/Data/Clean_CPI_Data.csv")

### 2024-01-23 Outputs Folder/paper_final.qmd
Fixed typos and updated appendix

--- 2024-01-23 Outputs Folder/paper_final.qmd
@@
 This report conducts an in-depth analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto].
-The author use the open-source statistical programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details.
+The author uses the open-source statistical programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details.
 In addition, this report further examines how economic changes indicated by external factors such as the Consumer Price Index (CPI) are related to changes in public health standards within the food industry.

### 2024-01-24 Outputs Folder/paper_final.qmd
Fixed typos and updated appendix

--- 2024-01-23 Outputs Folder/paper_final.qmd
@@
 #| warning: false
+# Load necessary libraries
+library(readr)
 library(dplyr)
-library(magrittr)  
+library(knitr)
 library(kableExtra)
-library(readr)
-library(tidyr)
-library(knitr)
 
 file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"
+
 dataset <- read_csv(file_path)
 
-processed_dataset <- dplyr::select(dataset, 
-                                   Inspection_Dates = `Inspection Date`, 
-                                   Establishment_Type = `Establishment Type`,
+selected_dataset <- dataset %>%
+  select(Establishment_Type = `Establishment Type`,
          Establishment_Status = `Establishment Status`,
-                                   Severity,
-                                   Action,
-                                   Min_Inspections_Per_Year = `Min. Inspections Per Year`)
+         Min_Inspections_Per_Year = `Min. Inspections Per Year`,
+         Severity) %>%
+  head(20)
 
-summary_table <- dplyr::group_by(processed_dataset, Establishment_Type, Establishment_Status) %>%
-  dplyr::summarise(Total_Count = n(), 
-                   Avg_Min_Inspections = mean(Min_Inspections_Per_Year, na.rm = TRUE)) %>%
-  dplyr::arrange(Establishment_Type, Establishment_Status)
-
-#knitr::kable(summary_table)
-
-kable(head(summary_table,20), format = "latex", longtable = TRUE, booktabs = TRUE) %>%
+kable(selected_dataset, format = "latex", longtable = TRUE, booktabs = TRUE) %>%
   kable_styling(latex_options = c("striped", "repeat_header"))

### 2024-01-24 Outputs Folder/paper_final.qmd
Fixed typos and updated appendix

@@
 
-file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"
+file_path <- "Inputs Folder/Data/Clean_Dinesafe_Data.csv"
 

### 2024-01-25 Scripts Folder/scripts/00-simulate_data.R
Delete Scripts Folder/scripts directory

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates whether the code and environment can run the required graphs well
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-#### Simulate data ####
-
-# Create a simulated for Dinesafe data set
-library(tidyverse)
-
-set.seed(123)
-simulated_dataset <- tibble(
-  `Establishment ID` = 1:100,
-  `Establishment Name` = paste("Establishment", 1:100),
-  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
-  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
-  `Infractions` = sample(0:5, 100, replace = TRUE)
-)
-
-# Create a simulated for Toronto's Dashboard Key Indicators data set
-set.seed(123)
-
-cpi_data <- tibble(
-  Year = 2010:2023, 
-  Average_CPI = runif(length(Year), min = 0, max = 5) 
-)
-

### 2024-01-25 Scripts Folder/scripts/02-data_cleaning.R

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: Clean Dinesafe and CPI dataset
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean data ####
-raw_data <- read.csv("Inputs Folder/Data/raw_data.csv")
-
-cleaned_data <- dataset %>%
-  filter(!is.na(`Establishment Name`)) %>%
-  mutate(new_column = tolower(`Establishment Name`))
-
-#### Save data ####
-write.csv(dataset, "Inputs Folder/Data/Clean_Dinesafe_Data.csv")
-
-#### Clean data ####
-raw_data2 <- read.csv("Inputs Folder/Data/raw_data2.csv")
-
-cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
-
-cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
-
-#### Save data ####
-write.csv(cpi_data, "Inputs Folder/Data/Clean_CPI_Data.csv")

### 2024-01-25 Scripts_Folder/scripts/00-simulate_data.R
Updated Boolean Test

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Simulates whether the code and environment can run the required graphs well
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+#### Simulate data ####
+
+# Create a simulated for Dinesafe data set
+library(tidyverse)
+
+set.seed(123)
+simulated_dataset <- tibble(
+  `Establishment ID` = 1:100,
+  `Establishment Name` = paste("Establishment", 1:100),
+  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
+  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
+  `Infractions` = sample(0:5, 100, replace = TRUE)
+)
+
+# Create a simulated for Toronto's Dashboard Key Indicators data set
+set.seed(123)
+
+cpi_data <- tibble(
+  Year = 2010:2023, 
+  Average_CPI = runif(length(Year), min = 0, max = 5) 
+)
+

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 1, 3 graphs

--- 2024-01-24 Outputs Folder/paper_final.qmd
@@
 The report extracted the complete data through R and classified the number of institutions by Establishment Status in the Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown.
-Taken together, the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards.
+Taken together, the count for the immediate "Pass" category is significantly higher with number of 76436, indicated by the higher bar on the right, indicating that more businesses are fully compliant with food safety inspection standards than those 369 businesses receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards.
 This visual demonstrates that most businesses inspected adhere to the required health and safety guidelines.
@@
 library(tidyverse)
+library(readr)
 
@@
   group_by(`Establishment Status`) %>%
-  summarize(Count = n())
+  summarize(Count = n(), .groups = 'drop') # It's good practice to drop groups after summarizing
 
@@
   geom_bar(stat = "identity") +
+  geom_text(aes(label = Count), vjust = -0.3, size = 3.5) + # Adds text on top of the bars
   theme_minimal() +
@@
        x = "Establishment Status", y = "Count")
+
 ```
@@
 ## Actions from Infraction Companies
-To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a pie chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
-Please note that the coding of this chart refers to the "Pie Chart customization" website [@rggpiechart].
+To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a bar chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
+Please note that the coding of this chart refers to the "Chart customization" website [@rggpiechart].
 The chart is dominated by the green proportion, representing the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action Toronto Public Health considers; instead, businesses are allowed to make amends within a certain period.
@@
 #| warning: false
-library(ggplot2)
-library(dplyr)
-library(lubridate)
+library(tidyverse)
+library(readr) # Make sure to load readr for read_csv()
 
@@
 
-ggplot(minor_severity_actions, aes(x = "", y = n, fill = Action)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar(theta = "y") + # This creates a pie chart, remove this line for a stacked bar chart
+ggplot(minor_severity_actions, aes(x = Action, y = n, fill = Action)) +
+  geom_bar(stat = "identity") +
   theme_minimal() +
   labs(
-       x = "Action", y = "Count")
+       x = "Action", y = "Count") +
+  coord_flip() # Flip the coordinates to make horizontal bars
 

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 3 graph

@@
   - Tianen (Evan) Hao
-thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git"
+thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Toronto_Dinesafe_DataAnalysis.git"
 date: January 21, 2024
@@
 Each category is divided into four violation severity levels: Minor, Significance, Crucial and Not Applicable (NA), which are further divided.
-This report referred to the relevant codes of the @rstats_top50ggplot2 website when writing the code for this bar chart.
+This report referred to the relevant codes of the "Top 50 Ggplot2 Visualization" [@rstats_top50ggplot2] website when writing the code for this bar chart.
 
@@
 To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a bar chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
-Please note that the coding of this chart refers to the "Chart customization" website [@rggpiechart].
+Please note that the coding of this chart refers to the "Top 50 Ggplot2 Visualization" [@rstats_top50ggplot2] website.
 The chart is dominated by the green proportion, representing the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action Toronto Public Health considers; instead, businesses are allowed to make amends within a certain period.

### 2024-01-25 Scripts_Folder/scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -10,21 +10,21 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read.csv("Inputs Folder/Data/raw_data.csv")
+raw_data <- read.csv("Inputs_Folder/Data/raw_data.csv")
 
 cleaned_data <- dataset %>%
   filter(!is.na(`Establishment Name`)) %>%
   mutate(new_column = tolower(`Establishment Name`))
 
 #### Save data ####
-write.csv(dataset, "Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+write.csv(dataset, "Inputs_Folder/Data/Clean_Dinesafe_Data.csv")
 
 #### Clean data ####
-raw_data2 <- read.csv("Inputs Folder/Data/raw_data2.csv")
+raw_data2 <- read.csv("Inputs_Folder/Data/raw_data2.csv")
 
 cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
 
 cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
 
 #### Save data ####
-write.csv(cpi_data, "Inputs Folder/Data/Clean_CPI_Data.csv")
+write.csv(cpi_data, "Inputs_Folder/Data/Clean_CPI_Data.csv")

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 3

@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")
+dataset <- read_csv("Inputs_Folder/Data/Clean_Dinesafe_Data.csv")
 
@@
 
-cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv") %>%
+cpi_data <- read_csv("Inputs_Folder/Data/Clean_CPI_Data.csv") %>%
   filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
@@
 
-dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
+dataset <- read_csv("Inputs_Folder/Data/Clean_Dinesafe_Data.csv") %>%
   mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
@@
 
-file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"
+file_path <- "Inputs_Folder/Data/Clean_Dinesafe_Data.csv"
 
@@
 
-file_path <- "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv"
+file_path <- "Inputs_Folder/Data/Clean_CPI_Data.csv"
 

### 2024-01-25 Scripts_Folder/scripts/02-data_cleaning.R
Delete Scripts_Folder/scripts directory

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: Clean Dinesafe and CPI dataset
-# Author: Tianen (Evan) Hao
-# Date: 14 January 2024
-# Contact: evan4830@gmail.com
-# License: MIT
-# Pre-requisites: R 4.3.2
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean data ####
-raw_data <- read.csv("Inputs_Folder/Data/raw_data.csv")
-
-cleaned_data <- dataset %>%
-  filter(!is.na(`Establishment Name`)) %>%
-  mutate(new_column = tolower(`Establishment Name`))
-
-#### Save data ####
-write.csv(dataset, "Inputs_Folder/Data/Clean_Dinesafe_Data.csv")
-
-#### Clean data ####
-raw_data2 <- read.csv("Inputs_Folder/Data/raw_data2.csv")
-
-cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
-
-cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
-
-#### Save data ####
-write.csv(cpi_data, "Inputs_Folder/Data/Clean_CPI_Data.csv")

### 2024-01-25 Scripts_Folder/scripts/01-download_data.R
Updated Boolean test

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Download and read Dinesafe and CPI data from OpenDataToronto
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+library(opendatatoronto)
+
+#### Download data ####
+# get package
+package <- show_package("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
+package
+
+# get all resources for this package
+resources <- list_package_resources("ea1d6e57-87af-4e23-b722-6c1f5aa18a8d")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+dataset <- filter(datastore_resources, row_number()==1) %>% get_resource()
+
+#### Save data ####
+write.csv(dataset, "raw_data.csv") 
+
+#### Download data 2####
+
+# Get package
+package <- show_package("c6d64e9b-f85a-4084-be14-60cf18509203")
+# It's a good practice to check the package details, but you can also comment it out if not needed
+# package
+
+# Get all resources for this package
+resources <- list_package_resources("c6d64e9b-f85a-4084-be14-60cf18509203")
+
+# Identify datastore resources; Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# Load the first datastore resource as a sample
+# Using `slice()` instead of `filter()` to select the first resource
+data <- slice(datastore_resources, 1) %>% get_resource()
+
+# View the first few rows of the dataset
+head(data)
+         
+#### Save data 2####
+write.csv(cpi_data, "raw_data2.csv") 

### 2024-01-25 Scripts_Folder/scripts/02-data_cleaning.R

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: Clean Dinesafe and CPI dataset
+# Author: Tianen (Evan) Hao
+# Date: 14 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read.csv("Inputs_Folder/Data/raw_data.csv")
+
+cleaned_data <- dataset %>%
+  filter(!is.na(`Establishment Name`)) %>%
+  mutate(new_column = tolower(`Establishment Name`))
+
+#### Save data ####
+write.csv(dataset, "Inputs_Folder/Data/Clean_Dinesafe_Data.csv")
+
+#### Clean data ####
+raw_data2 <- read.csv("Inputs_Folder/Data/raw_data2.csv")
+
+cpi_data <- filter(data, measure_name == "CPI - Consumer Price Index - Toronto (% Change)")
+
+cpi_data_renamed <- rename(cpi_data, CPI = measure_value)
+
+#### Save data ####
+write.csv(cpi_data, "Inputs_Folder/Data/Clean_CPI_Data.csv")

### 2024-01-25 Scripts_Folder/scripts/03-test_data.R

@@ -1,0 +1,106 @@
+#### Preamble ####
+# Purpose: Tests whether dataset can operate normally
+# Author: Tianen (Evan) Hao
+# Date: 15 January 2024
+# Contact: evan4830@gmail.com
+# License: MIT
+# Pre-requisites: R 4.3.2
+
+
+#### Test dinesafe data ####
+library(tidyverse)
+
+# Use the simulated DineSafe data set
+set.seed(123)
+simulated_dataset <- tibble(
+  `Establishment ID` = 1:100,
+  `Establishment Name` = paste("Establishment", 1:100),
+  `Establishment Status` = sample(c("Pass", "Conditional Pass", "Fail"), 100, replace = TRUE),
+  `Inspection Date` = sample(seq(as.Date('2021-01-01'), as.Date('2021-12-31'), by="day"), 100),
+  `Infractions` = sample(0:5, 100, replace = TRUE)
+)
+
+# Test if 'Establishment Status' contains specific categories
+all(c("Pass", "Conditional Pass", "Fail") %in% simulated_dataset$`Establishment Status`)
+
+# Test the range of dates in 'Inspection Date'
+simulated_dataset$`Inspection Date` |> min() >= as.Date('2021-01-01')
+simulated_dataset$`Inspection Date` |> max() <= as.Date('2021-12-31')
+
+# Test the range of values in 'Infractions'
+simulated_dataset$Infractions |> min() >= 0
+simulated_dataset$Infractions |> max() <= 5
+
+# Test the class of 'Establishment ID'
+simulated_dataset$`Establishment ID` |> class() == "integer"
+
+# Some extra test
+status_count <- simulated_dataset %>%
+  group_by(`Establishment Status`) %>%
+  summarize(Count = n())
+
+ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  theme_minimal() +
+  labs(title = "Number of Establishments by Status",
+       x = "Establishment Status", y = "Count")
+
+library(dplyr)
+
+inspection_summary <- simulated_dataset %>%
+  group_by(`Establishment Status`) %>%
+  summarise(Count = n()) %>%
+  mutate(Percentage = (Count / sum(Count)) * 100)
+
+print(inspection_summary)
+
+
+#### Test Toronto's Dashboard Key Indicators data set####
+
[…]

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 3

--- 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
@@
 abstract: "Nowadays, as the market economy continues to thrive, people have more diverse choices of edible products and are also more aware of their potential health risks.
-This report uses the open-source statistical programming language R to analyze a data set of food safety inspections at Toronto dining establishments and further uses additional data sets to correlate food safety trends with an economic indicator such as the Consumer Price Index (CPI).
+This report analyzes a data set of food safety inspections at Toronto dining establishments and further uses additional data sets to correlate food safety trends with an economic indicator such as the Consumer Price Index (CPI).
 The analysis shows a limited reduction in sanitation infractions; this potentially means food safety compliance is improving in the City of Toronto.
-Additionally, the analysis suggests some potential links between economic downturns and increases in health infractions."
+Additionally, the analysis suggests some potential links between economic downturns and increases in health infractions; this initially establishes a correlation between economic contraction and increases in food safety infractions, with the aim of further elucidating these relationships and their implications for public policy."
 format: pdf
@@
 
-This report conducts an in-depth analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto].
+This report conducts an analysis of Dinesafe data provided by the City of Toronto's open data website OpenDataToronto [@opendatatoronto].
 The author uses the open-source statistical programming language R [@citeR] to display trends in the data, focusing on key indicators such as inspection pass rate and changes in infraction details.
 In addition, this report further examines how economic changes indicated by external factors such as the Consumer Price Index (CPI) are related to changes in public health standards within the food industry.
-Through analysis, results found that restaurants are more likely to violate food safety regulations, and there is a correlation between CPI fluctuations and infraction behaviours.
+Through analysis, results found that restaurants are more likely to violate food safety regulations, and there is a possible correlation between CPI fluctuations and infraction behaviours.
 The trend shows that the number of infraction behaviours will increase when CPI decreases.
@@
 Finally, minimum inspections per year serve as a reference for analyzing food safety status and corporate behaviour.
-Considering the hugeness of the dataset data, please refer to \ref{app:clean-dinesafe-dataset}.
+Considering the hugeness of the dataset data, there is a preview of the data is included in Appendix A; please refer to \ref{app:clean-dinesafe-dataset}.
 ```{r}
@@
 This dataset contains many economic-related indicators that the Toronto government is open to the public; the author selected the relevant period between January 2022 and January 2023 of the Consumer Price Index (CPI), including numerical values and years that are consistent with Dinesafe for analysis.
-Considering the hugeness of the dataset data, please refer to \ref{app:clean-cpi-dataset}.
+Considering the hugeness of the dataset, a preview of the data is included in Appendix B; please refer to \ref{app:clean-cpi-dataset}.
 
@@
 The report extracted the complete data through R and classified the number of institutions by Establishment Status in the Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown.
-Taken together, the count for the immediate "Pass" category is significantly higher with number of 76436, indicated by the higher bar on the right, indicating that more businesses are fully compliant with food safety inspection standards than those 369 businesses receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards.
+Taken together, based on the large sample size, the difference is significant; the count for the immediate "Pass" category is significantly higher with the number of 76436, indicated by the higher bar on the right, indicating that more businesses are fully compliant with food safety inspection standards than those 369 businesses receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards.
 This visual demonstrates that most businesses inspected adhere to the required health and safety guidelines.
@@
 
-Significantly fewer infractions were labelled major than minor ones, which may indicate that more serious health risks are less likely to occur.
+Significantly fewer infractions were labelled major than minor ones as [@tbl-total_count_severity_levels] shown, which may indicate that more serious health risks are less likely to occur.
 Nonetheless, the presence of these infractions indicates the need for improvements to ensure public health.
@@
 ## Actions from Infraction Companies
-To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a bar chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions].
+To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a bar chart to more clearly visualize the data in [@fig-distribution_minor_severity_infractions].
 Please note that the coding of this chart refers to the "Top 50 Ggplot2 Visualization" [@rstats_top50ggplot2] website.

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 3

--- 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
@@
 
+# Future Expectations
+Although this report does not provide an in-depth analysis of Toronto's food safety situation and related economic indicators, the results of these studies provide some practical assistance for several future research avenues.
+Future investigations into food safety infraction behaviours' geographic and demographic distribution could reveal patterns and underlying inequalities across Toronto.
+Additionally, examining differences in infraction behaviour between different types of cuisine may provide insight into specific risk factors associated with food preparation practices.
+Finally, further exploring the relationship between economic recession and food safety compliance through comprehensive economic indicators rather than just CPI can provide a more detailed understanding of how economic pressure affects public health standards.
+The report believes these future studies will contribute to a more comprehensive understanding of food security in urban settings and inform more effective public health policies and interventions.
+
 \appendix
@@
 Considering the huge content of the data set, this appendix contains a preview of the cleaned data set that has been used for the research.
-Original data is avaliable at https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git
+Original data is avaliable at https://github.com/ShadyEvan4830/Toronto_Dinesafe_DataAnalysis.git
 ```{r}
@@
 Considering the huge content of the data set, this appendix contains a preview of the cleaned data set that has been used for the research.
-Original data is avaliable at https://github.com/ShadyEvan4830/Tonronto_Dinesafe_DataAnalysis.git
+Original data is avaliable at https://github.com/ShadyEvan4830/Toronto_Dinesafe_DataAnalysis.git
 
